---
title: "Genomic landscape of transposable elements and structural variation using long reads whole genome sequencing"
author: "Yocelyn T. Gutiérrez-Guerero"
output: pdf_document
fontsize: 12pt
header-includes:
 \usepackage[default]{sourcesanspro}
 \usepackage[T1]{fontenc}
 \usepackage{fancyhdr}
 \usepackage{helvet}
 \renewcommand{\familydefault}{\sfdefault}
mainfont: SourceSansPro
---
# Index

## Download PacBio reads

## PacBio quality check

## Convert BAM to Fastq and Adapter removal

## Genome Assembly

#### [Flye](https://github.com/fenderglass/Flye) assembler

#### [Hifiasm](https://github.com/chhylp123/hifiasm) assembler



Installation requirements

-PacBio data requires to be processed using the SMRTlink software or PacBio tools that are distributed via Bioconda [pbbioconda](https://github.com/PacificBiosciences/pbbioconda)


##Download PacBio reads

```{bash, eval=FALSE}

fasterq-dump SRA_Accession -F fastq -o Out_pb -e 2

```


Converting BAM files to Fastq format and removing adapters using [HiFiAdapterFilt](https://github.com/sheinasim/HiFiAdapterFilt)
*Dependencies*: python, blast, bamtools

```{bash, eval=FALSE, tidy=TRUE}

pbadapterfilt.sh  -t 20 -o ~/PacBio_filter_fastq/

```


Genome assembly with Flye, including a correction with Minimap2

```{bash eval=FALSE, tidy=TRUE}

flye --pacbio-hifi ID_raw_pacbioRev.fastq.gz --genome-size 3.1g 
-o PacBio_assembly/ -t 40 -i 2 --scaffold --asm-coverage 20

```

Genome assembly with [Hifiasm](https://github.com/chhylp123/hifiasm) and genome quality using quast

```{bash, eval=FALSE}

mkdir hifiasm_ID
cd hifiasm_ID/
ln -s ~/PacBio_fastq/ID_raw_pacbioRev.fastq.gz .
hifiasm -o ID_hifiasm.asm -t 30 ID_raw_pacbioRev.fastq.gz -f38

awk '/^S/{print ">"$2;print $3}' ID_hifiasm.asm.bp.p_ctg.gfa 
> ID_hifiasm_ctg.fasta

quast.py ID_hifiasm_ctg.fasta -o ID_hifiasm_stats.txt

```

Scaffolding and polishing with Gapless

```{bash, eval=FALSE}

gapless.py split -o gapless_split.fa ID_hifiasm_ctg.fasta
minimap2 -t 30 -DP -k19 -w19 -m200 gapless_split.fa gapless_split.fa
> gapless_split_repeats.paf

minimap2 -t 30 -x map-hifi -c -N 5 --secondary=no gapless_split.fa
ID.hifi_reads.default.filt.fastq.gz > gapless_reads.paf

gapless.py scaffold -p gapless -s gapless_stats.pdf gapless_split.fa
gapless_reads.paf gapless_split_repeats.paf

minimap2 -t 30 -x map-hifi  <(seqtk subseq ID.hifi_reads.default.filt.fastq.gz
gapless_extending_reads.lst) <(seqtk subseq ID.hifi_reads.default.filt.fastq.gz
gapless_extending_reads.lst) > gapless_extending_reads.paf

gapless.py extend -p gapless gapless_extending_reads.paf

seqtk subseq ID.hifi_reads.default.filt.fastq.gz gapless_used_reads.lst
> temp_finish.fastq

gapless.py finish -o gapless_raw.fa -H 0 -s 
gapless_extended_scaffold_paths.csv -p gapless_polishing.csv 
gapless_split.fa temp_finish.fastq

minimap2 -t 30 -x map-hifi gapless_raw.fa ID.hifi_reads.default.filt.fastq.gz
> gapless_consensus.paf

rm gapless_split.fa gapless_split_repeats.paf gapless_reads.paf

gapless_stats.pdf gapless_scaffold_paths.csv gapless_extensions.csv 

gapless_extending_reads.lst gapless_polishing.csv 
gapless_extending_reads.paf gapless_extended_scaffold_paths.csv 
gapless_used_reads.lst 
gapless_extended_polishing.csv gapless_raw_polishing.paf temp_finish.fastq
```

Polishing with Minimap2, pbmm2, samtools and Racon 
```{bash, eval=FALSE}

pbmm2 index ID_hifiasm_ctg.fasta ID_hifiasm_ctg.fasta.mmi
pbmm2 align ID_hifiasm_ctg.fasta ID.hifi_reads.default.filt.fastq.gz 
ID_align.bam --sort -j 20

samtools view -h ID_align.bam > ID_align.sam

racon id-t 20 ID.hifi_reads.default.filt.fastq.gz ID_align.sam 
ID_hifiasm_ctg.fasta > ID_assembly_racon.fasta

rm  ID_align.bam ID_align.sam ID_align.bai

##Evaluate genome assembly

quast.py ID_assembly_racon.fasta -o ID_assembly_stats


```


###Busco

```{r, eval=FALSE}

library(ggplot2)
library(webr)
library(dplyr)

bdata<-data.frame(Category= c("Single-copy","Duplicated","Fragmented",
                              "Missing") ,Value=c(8444,158,108,516))


dbusco = bdata %>% group_by(Category, Class) %>% summarise(n = sum(Value))

pie_chart <- ggplot(bdata, aes(x = "", y = Value, fill = Category)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar(theta = "y") +
    geom_text(aes(x = 1.6, label = paste(round(Value/sum(Value) * 100), "%")), 
              position = position_stack(vjust = 0.5)) +
    scale_fill_manual(values=maize_pal("FloweringTime")) +
    theme_void()

donut_chart <- pie_chart + 
    geom_bar(data = bdata, aes(x = "", y = Value), stat = "identity", 
             width = 0.5, fill = "white") +
    geom_text(data = bdata, aes(label = paste(round(Value/sum(Value) * 100),
                                              "%")), 
              position = position_stack(vjust = 0.5), color = "white") +
    theme_void()

grid.arrange(pie_chart, donut_chart, ncol = 2, widths = c(3, 3))


```

### Ragtag:

Connect contigs into chromosomes using homology (if a reference genome is available)

```{bash, eval=FALSE}

ragtag.py scaffold ReferenceGenome_genomic.fna ID_assembly_racon.fasta -u 
-t 30
```

## Mitogenome identification and circulization

Using the House mouse mitogenome from the genome reference

```{bash, eval=FALSE}

python mitohifi.py -c ID_assembly_racon.fasta -f RefGenone_Mit.fasta 
-g RefGenome_Mit.gbk -t 10 -a animal -o 2

##Verification

cat NC_005089.1_RagTag.fasta RefGenone_Mit.fasta > seq_NC_005089.fasta

mafft --anysymbol --parttree --retree 1 --thread 5 seq_NC_005089.fasta
> align_NC_005089.fasta

```

## Masking mitogenome

```{bash, eval=FALSE}

##create a bed file, indicating the coordinates:
##ptg000200l	521	16831

bedtools maskfasta -fi ID_ragtag_scaffold.fasta -bed mito.bed 
-fo ID_mitomask.fasta

```

## Identify TE

### Repeat Modeler

```{bash, eval=FALSE}

BuildDatabase -name ID_db -engine ncbi ID_ragtag_scaffold.fasta

RepeatModeler -engine ncbi -pa 10 -database ID_db

```

## MCHelper

```{r, eval=FALSE}

cat ~/busco/mammalia_odb10/hmms/*.hmm > mammals_odb10.hmm

## Automatic Curation
python3 MCHelper.py -r A -t 20 -l ID_db-families.fa -o MCHelper_out/ -g
~/hifiasm/ID_assembly_racon.fasta --input_type fasta -b mammals_odb10.hmm -a F

## Manual Curation
python3 MCHelper.py -r M -l ~/MCHelper/Curation/Pop1_3id_curated_NR.fa -o
~/MCHelper/Curation/Pop1_3id_curated --input_type fasta -g
~/PacBioData/Final_assembly/ID_assembly_racon.fasta -t 30


```

## Repeat Masker Annotation

```{bash, eval=FALSE}

RepeatMasker -pa 50 -lib MCHelperDfam_TElibrary_NR_rename.fasta -x 
-gff ID_assembly_racon.fasta

```

